\section{Results}
\label{results}
\todo{pick results}
In this section I will show and discuss the most important results I have found while testing the Geometrical and Numerical algorithms. 

I have only included the most important graphs based on the data sets, as well as a table detailing the speed and spread of the 2 different algorithms, since the generated data set is too large to practical include in this report.

The complete data-sets can be found in this projects github repository at \underline{https://github.com/kjmikkel/LonelyRunner}

When a graph in the following contains a hyphenated ``Random'' in its title, it means that while both the Geometrical and the Numerical algorithm has been given the same input, the order of the input for the Numerical algorithm has been randomized. 


\subsection{Graphs}

\subsubsection{Primes-Random}
%\graph{Primes-Random}{500}{Geo}{}
%\graph{Primes-Random}{1000}{Geo}{}

\subsubsection{Sequential Numbers}
Figure \ref{Sequential_100_Num} and \ref{Sequential_100_Geo} is the same figure, just with different ranges on the y-axis.

\graph{Sequential}{100}{Num}{The time anomaly for the Numerical algorithm is very clear}
\graph{Sequential}{100}{Geo}{We can see, that once the anomaly is over, the Numerical algorithm is clearly faster than the Geometrical algorithm}

The above pattern is one that appears every time the input is Sequential, but only for the first configuration.

\subsubsection{Sequential-Random}
\graph{Sequential-Random}{100}{Num}{Please note the very large Error bar, which is not present in figure \ref{Sequential_100_Num}. This is because of randomized order of the input}


%\graph{Sequential-Random}{1000}{Num}{}
%\graph{Sequential-Random}{1000}{Geo}{}
\graph{Sequential-Random}{2000}{Num}{This graph is very interesting because it shows us, that even when the order has been permuted, there is still a high chance of having the Numerical run for a very long time for the first case. In this case it is around 7 hours. It is also very interesting to note that the Error Bar is very small, meaning that the permutation of the order does not guarantee that a fast solution can be found}
%\graph{Sequential-Random}{500}{Num}{}

\subsubsection{Random}
%\graph{Random}{500}{Geo}{}
%\graph{Random}{2000}{Geo}{}
%\graph{Random}{4000}{Geo}{}

\subsection{Sequential results}
\subsubsection{Random Sequential Results}

\input{./data/tables/sequential-random_speed_table.tex}
 
\input{./data/tables/sequential-random_spread_table.tex}

\subsubsection{Non-Random Sequential Results}

\input{./data/tables/sequential-normal_speed_table.tex}
 
\input{./data/tables/sequential-normal_spread_table.tex}

\subsubsection{Combined Sequential Results}

\input{./data/tables/sequential_speed_table.tex}
 
\input{./data/tables/sequential_spread_table.tex}

From the above it is clear that, not only is the Geometrical algorithm more stable than the Numerical algorithm, but I would also argue that, for this instance, it is the preferable algorithm. Even if we were to use somehow could exclude the extreme cases, we would only save 3 minutes by using the Numerical algorithm. Clearly it is not worth risking about 12 hours of run-time for that. 

\subsection{Random results}
\label{random_results}
\subsubsection{Non-sorted Random Results}
\input{./data/tables/random-random_speed_table.tex}

This result is actually rather surprising. Since the random numbers tend to be rather large (since they lie between 1 and $2^{32} - 1$), I would have expected the Geometrical solution to be slower, since more Event Points would have had to be created, inserted and removed from the Event Queue, before any of the Runners could enter the Zone. The fact that there are no calculations by the Numerical algorithm that took over 1 second is not that surprising, given that completely random speeds were being used for the tested configurations.\\

\input{./data/tables/random-random_spread_table.tex}

Again it is a surprise that the spread for the Geometrical algorithm is less than for the Numerical in the best case, and not that much worse in the worst case.

\subsubsection{Sorted Random Results}


\subsubsection{Combined Results for the Random Numbers}
\input{./data/tables/random_speed_table.tex}

\input{./data/tables/random_spread_table.tex}


\subsection{Prime Results}
\label{prime_results}

\subsubsection{Random Prime Results}
\label{random_prime_results}

\input{./data/tables/prime-random_speed_table.tex}

\input{./data/tables/prime-random_spread_table.tex}

\subsubsection{Non-Random Prime Results}

\input{./data/tables/prime-normal_speed_table.tex}

It is here interesting to note that in this case the Numerical algorithm is slightly slower than in the last section, where the runner speeds had been randomized, clearly showing that the order of the inputs matter.\\

\input{./data/tables/prime-normal_spread_table.tex}

Comparing the spread in this subsection with the one in Section \ref{random_prime_results} clearly serves to illustrate that randomising the input has a clear effect on the time needed for the Numerical algorithm to find a solution.

\subsubsection{Combined Results for the Primes Numbers} 

\input{./data/tables/prime_speed_table.tex}

\input{./data/tables/prime_spread_table.tex}

Based on the above 2 tables, it seems clear to me that the Prime numbers contain properties that makes the Numerical algorithm more suitable. It is also clear that none of the algorithms is able to find a solution right away, but rather has to work on it. 

\input{./data/tables/prime_speed_table.tex}

\input{./data/tables/prime_spread_table.tex}

\subsection{Total Results}
\subsubsection{Random Total Results}

\input{./data/tables/total-random_speed_table.tex}

\input{./data/tables/total-random_spread_table.tex}

\subsubsection{Non-Random Total results}

\input{./data/tables/total-normal_speed_table.tex}

\input{./data/tables/total-normal_spread_table.tex}

\subsubsection{Combined Results} 

\input{./data/tables/total_speed_table.tex}

\input{./data/tables/total_spread_table.tex}

\subsection{Speed}

From the above graphs it would seem tempting to assume that the Numerical algorithm was dramatically better than the Geometrical, even if we included the extreme times reported, however a closer look shows that the truth is more complicated than that. The below table and calculations are based on all the reported speeds:

%\input{./data/tables/speed_table.tex}

While it from the graphs would seem clear that the Numerical algorithm is faster for most of the instances. 
I believe the reason for this to belong to the fundamental differences of the 2 algorithms. While both algorithms are based on exploring all possible candidates times for a time where equation \eqaref{eqa:lonelyRunner} is true, the Geometrical algorithm works by sweeping through the entire first round of the final runner\footnote{The runner with speed 1}, while the Numerical algorithm often is less methodical about it, depending on the 2 selected runner speeds, trying time points against equation \eqaref{eqa:lonelyRunner}. In practice it is clear this is a better approach than the Geometrical algorithm.
 
It however also clear that for some configurations (even when accounting for the spread), finding a time where equation \eqaref{eqa:lonelyRunner} is true, takes much more time than most of the other configurations for that number of runners. Two good examples of this are \figureref{Primes_1000_Geo} and \figureref{Sequential-Random_1000_Geo}. The best explanation I have been able to come up with for cases like the first is that equation \eqaref{eqa:lonelyRunner} depends not so much on the size of the numbers, but rather on the on the number of the runners and the interaction of the runners speeds. I believe cases like the second are caused by the input based run-time of the Numerical algorithm - if the order of the runner speeds are in such a way that a good part of list has to be searched through, then the Numerical algorithms run-time of $O(k * n^3)$ (as proved back in Section \ref{proof_num} on page \pageref{proof_num}) begins to show itself.  

\subsection{Spread}
%\input{./data/tables/spread_table.tex}



\subsection{Conclusion}

From the above it is clear that the Numerical algorithm is faster than the Geometrical. It is also clear from the graphs that neither of the algorithms are entirely stable, and for some configurations have both a very large execution time spread. As I have not been able to find a pattern for the configurations that have produced some of these abnormalities, I can 
